Approximation as Differentiation

### 2025-06-09 main/first-order.tex
Minor clean-up.

We now consider how to adapt the CHAD framework~\cite{vákár22,nunes2023} to interpret programs as functions
along with their forward and backward approximation maps. Since the appropriate lattice of approximation
typically depends on the value being approximated, we will want to interpret a type $\tau$ as a pair $(X,
\partial X)$ where $X$ provides the conventional interpretation as a set and $\partial X$ provides an
appropriate lattice of approximations $\partial X(x)$ for every point $x \in X$. An expression $e$ of type
$\sigma$ with a single free variable of type $\tau$ (supposing $\tau$ and $\sigma$ denote $X$ and $Y$ in the
conventional semantics) must then also be interpreted by a pair $(f, \partial f)$, where $f$ provides the
conventional interpretation of $e$ as a function $X \to Y$, and $\partial f$ is a family of Galois connections
$\partial f(x): \partial X(x) \to \partial Y(f(x))$ for every $x \in X$, taking approximants to approximants.

@@

\subsection{Category of Families}
\label{sec:Fam}

Now we formalise $X$-indexed families as functor categories $\Func{X}{\cat{C}}$ and consider the specific case
of the Grothendieck construction for set-indexed categories $\Func{-}{\cat{C}}: \Set^\op \to \Cat$.

\begin{definition}[Category of $X$-indexed families]
For any set $X$ (regarded as a discrete category) and any category $\cat{C}$, recall that the functor
category $\Func{X}{\cat{C}}$ has as objects all functors $F: X \to \cat{C}$, namely $X$-indexed families of
objects of $\cat{C}$, and as morphisms from $F \to G$, all families of morphisms $\eta(x): F(x) \to G(x)$ in
$\cat{C}$ for every $x \in X$, with naturality trivial because $X$ has only identity morphisms.
\end{definition}

>>>>

We now consider how to adapt the CHAD framework~\cite{vákár22,nunes2023} to interpret programs as functions
along with their forward and backward approximation maps. Since the appropriate lattice of approximation
typically depends on the value being approximated, we will want to interpret a type $\tau$ as a pair $(X,
\partial X)$ where $X$ provides the conventional interpretation as a set and $\partial X$ provides an
appropriate lattice of approximations $\partial X(x)$ for every point $x \in X$. An expression $e$ of type
$\sigma$ with a single free variable of type $\tau$ (supposing $\tau$ and $\sigma$ denote $X$ and $Y$ in the
conventional semantics) must then also be interpreted by a pair $(f, \partial f)$, with $f$ providing the
conventional interpretation as a function $X \to Y$, and $\partial f$ as a family of Galois connections
$\partial f(x): \partial X(x) \to \partial Y(f(x))$ taking approximants to approximants.

@@

\subsubsection{Category of Families}
\label{sec:Fam}

For CHAD we are interested in the specific case of the Grothendieck construction where $\cat{C} = \Set$. In
that case our indexed categories $T$ have the form $\Func{-}{\cat{C}}: \Set^\op \to \Cat$, associating to each
set $X$ the functor category $\Func{X}{\cat{C}}$ which has as objects all functors $F: X \to \cat{C}$, namely
$X$-indexed families of objects of $\cat{C}$, and as morphisms $F \to G$, all families of morphisms $\eta(x):
F(x) \to G(x)$ in $\cat{C}$, with naturality trivial because $X$ has only identity morphisms.
